---
title: "Notes"
output:
  html_document:
    include:
      before_body: header.html
---

# Tuesday

Read in simple csv file named `test.csv` that looks like this

```
  Year Species metric.tons
1 2018   Fish1           1
2 2019   Fish1           2
3 2018   Fish2           3
4 2019   Fish2           4
5 2018   Fish3           6
6 2019   Fish4          NA
```

```
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
save(test, file="test.RData")
```

What is a list?  It is a object with multiple other types of objects 'tied' together.

```{r}
a <- list(ar=c(.8,.3), name="model", years=1:10)
a
```

You can now reference the different objects in `a`.

```{r}
a$ar
a$name
```

----

Why is the `arima.sim` throwing an error sometimes?

# Monday

Please install the rmarkdown package.  I forgot to include that.

```
install.packages("rmarkdown")
```

----

If you do not have LaTeX installed, you can install the tinytex package.  Run these commands.  This will allow you to make PDF files from R Markdown files.  Some of you installed tinytex, but you need to run the 2nd command also to install LiveTeX.

```
install.packages('tinytex')
tinytex::install_tinytex()
```

----

If your get an error saying package or repository unavailable when you are trying to install packages, try the following:

* Make sure you are online.
* Run `chooseCRANmirror()` from the command line.  Pick any mirror.
* Try installing package again.  Repeat with a different mirror if it doesn't work.

-----

<hr>


<div style="text-align: center">
<i class="fas fa-cubes"></i> NOAA Fisheries - NW Fisheries Science Center <i class="fas fa-cubes"></i>
</div>
