---
title: "Notes"
output:
  html_document:
    include:
      before_body: header.html
---

# Thusday 

Seasonal data.  If you have seasonal data that you would like to analyze, please put in this format:

```
  Year   Month  metric.tons
 2018   1           1
 2018   2           2
 2018   3           3
 ...   
 2019   1           4
 2019   2           6
 2019   3          NA
```

Read in with this code:

```
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
save(test, file="test.RData")
```

# Wednesday

To save data in R to a .csv file use the following:

```
write.csv(landings, file="test.csv", row.names=FALSE, quote=FALSE)
```

This works for data.frames, like landings, and also matrices:

```
mat <- matrix(rnorm(10),5,2)
colnames(mat) <- c("number 1", "number 2")
write.csv(landings, file="test.csv", row.names=FALSE, quote=FALSE)
```


# Tuesday

The bookdown-demo repository has been edited so that it does not try to make a PDF.  If you want that, there are instructions in the README.md file.

---

Read in simple csv file named ```test.csv``` that looks like this

```
  Year Species metric.tons
1 2018   Fish1           1
2 2019   Fish1           2
3 2018   Fish2           3
4 2019   Fish2           4
5 2018   Fish3           6
6 2019   Fish4          NA
```
with this code:

```
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
save(test, file="test.RData")
```

---

Read in a file where the data are in columns.  If your data (test.csv) look like this with each species (or site) across the columns:

```
Year,Anchovy,Sardine,Chub mackerel,Horse mackerel,Mackerel,Jack Mackerel
1964,5449.2,12984.4,1720.7,4022.4,NA,NA
1965,4263.5,10611.1,1278.5,4158.3,NA,NA
1966,5146.4,11437.8,802.6,3012.1,NA,NA
```

Use this code:

```
library(reshape2)
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
melt(test, id="Year", value.name="metric.tons", variable.name="Species")
save(test, file="test.RData")
```

If your data also have, say, a month (or qtr) column, use this code:

```
library(reshape2)
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
melt(test, id=c("Year","Month"), value.name="metric.tons", variable.name="Species")
save(test, file="test.RData")
```

----

If you have a response variable and multiple explanatory variables:

```
Year, Anchovy, SST,  Mackerel
1964, 5449.2,  24.4, 1720.7
1965, 4263.5,  30.1, 1278.5
1966, 5146.4,  23.8,  802.6
```

Use this code:

```
test <- read.csv("Data/test.csv", stringsAsFactors = FALSE)
save(test, file="test.RData")
```

Use this `lm()` model (or gam() etc):

```
fit <- lm(Anchovy ~ SST + Mackerel, data=test)
```

---

What is a list?  It is a object with multiple other types of objects 'tied' together.

```{r}
a <- list(ar=c(.8,.3), name="model", years=1:10)
a
```

You can now reference the different objects in `a`.

```{r}
a$ar
a$name
```

----

Why is the `arima.sim` throwing an error sometimes?  You must specify a stationary model and if you just randomly chose the $\beta$'s (AR parameters) and $\theta$'s (MA parameters) then you might be specifying a non-stationary model.

# Monday

Please install the rmarkdown package.  I forgot to include that.

```
install.packages("rmarkdown")
```

----

If you do not have LaTeX installed, you can install the tinytex package.  Run these commands.  This will allow you to make PDF files from R Markdown files.  Some of you installed tinytex, but you need to run the 2nd command also to install LiveTeX.

```
install.packages('tinytex')
tinytex::install_tinytex()
```

----

If your get an error saying package or repository unavailable when you are trying to install packages, try the following:

* Make sure you are online.
* Run `chooseCRANmirror()` from the command line.  Pick any mirror.
* Try installing package again.  Repeat with a different mirror if it doesn't work.

-----

<hr>


<div style="text-align: center">
<i class="fas fa-cubes"></i> NOAA Fisheries - NW Fisheries Science Center <i class="fas fa-cubes"></i>
</div>
