---
title: "Set-up"
output:
  html_document:
    toc: true
    include:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following will walk you through installing R, RStudio, and GitHub Desktop.  Though the instructions are long, it hopefully will not take you long to get set up.  You will need admin access on your laptop.

If you run into trouble, you can post in the comment section at the bottom.   Please, help each other out in the comments if you can solve someone's installation issues.  I will check in periodically and try to help too.

## Get an account on GitHub

If you don't already have an account on GitHub, go to [github.com](github.com) and click the "sign up" link near upper right 
of the page.  It is pretty self-explanatory.  Go ahead and get a **free** account.  

The free GitHub account does not have private repositories.  Everything is public and viewable.  You can upgrade (for money) and get private repositories.  Or _if you are a student_, you can get private repositories from GitHub for free.  Go to [https://education.github.com/pack](https://education.github.com/pack) to sign up for your free student pack.  You will need to upload proof that you are a student or faculty.

## Install GitHub Desktop (Windows and Mac users)

1. Go to [desktop.github.com](https://desktop.github.com/) and install.
2. Open GitHub Desktop, go to 'GitHub Desktop > Preferences' menu.  
3. Under 'Accounts', sign into your GitHub account.
3. Under 'Git', enter your name (or initials) and the email you used for your GitHub account.

Linux users can install Git and interact with GitHub though RStudio.  GitHub Desktop is not essential for the course, however it is a standard R Workflow that I will be demonstrating.

## Install R and RStudio

* **R:**  Make sure you have the latest version of R.  
Go to [https://cran.r-project.org/](https://cran.r-project.org/) and find the download link for your computer system.

* **RStudio:** Install the latest version of RStudio.  Get it from [https://www.rstudio.com/products/rstudio/download](https://www.rstudio.com/products/rstudio/download) and install the appropriate one for your OS.

## Set up RStudio to use Git

1. Open RStudio
2. Go to Global Options (from the Tools menu)
3. Click Git/SVN
4. Click Enable version control interface for RStudio projects

Linux users will need to install Git for this to work.  Windows/Mac users: Git comes with GitHub Desktop.

## Install R packages

From the command line in R Studio

1. **bookdown:** package.  Enter the following code at the command line (`>`).

```{r get-bd, eval=FALSE}
install.packages("devtools")  
devtools::install_github("rstudio/bookdown")
```

2. Install **other packages** that we are going to be need. I have created a package on GitHub that will install the packages you need.  This will probably take **awhile** as the first time you do it, many packages will be loaded.

```{r get-packages, eval=FALSE}
devtools::install_github("RVerse-Tutorials/RWorkflowsetup")
```

## Create a workshop folder

Create a folder/directory on your computer for the workshop materials.  You can create it anywhere you wish.  Name it `RWorkflow` just so we all use the same folder name for the workshop.

## Disqus forum

I have added a Disqus forum to the website.  [Create an account](https://disqus.com/) if you would like to be able to add comments/questions.  This will allow people to pose questions to the group and allow me (and others) to answer.

## Downloading repositories

**In case we do not have internet access, I will provide the files on a thumb drive also.**

You will need to fork (copy repositories from [RVerse-Tutorials](https://github.com/RVerse-Tutorials) for many of the labs. Please try to do this with a test repository that I have set up for you. 

1. Log into GitHub.
2. Go to [RVerse-Tutorials/Test](https://github.com/RVerse-Tutorials/Test).
3. Click 'Fork' in the upper right corner of the repository.  This will create a copy of Test in your GitHub account.
4. Open RStudio.  Click File > New Project > Version Control
5. Paste in the url of the repository you are copying---the **forked** repository in your GitHub account.  For example, `https://github.com/<youraccount>/Test`  You can leave project name blank.  Replace `<youraccount>` with your GitHub account.
6. You should now be able to see the files from Test in RStudio.  There is only one file, `README.md`.

[Watch a video that shows you how to do this](https://youtu.be/b6AaTHFIY7U)

## Packages with C++ code (Optional)

We will not be doing this in this short course, but often you will want to install packages with C++ code.  To do that, you will need Rtools (Windows) or Xcode (Mac)

* **Rtools** Windows users  [Rtools](https://cran.r-project.org/bin/windows/Rtools/) also so you can build packages with C++ code if needed.  See comments here about changing path. [Rtools](http://stat545.com/packages01_system-prep.html#windows-system-prep)

* **Xcode** Mac users Open terminal and type the following command `xcode-select --install`


